---
title: "R-intro "
output: pdf_document
indent: true
header-includes:
  # - \usepackage{titling}
  # - \preauthor{\begin{flushleft}}
  # - \postauthor{\end{flushleft}}
  # - \predate{\begin{flushleft}}
  # - \postdate{\end{flushleft}}
  - \usepackage{fancyhdr}
  - \pagestyle{fancy}
  - \fancypagestyle{plain}{\pagestyle{fancy}}
  - \headsep = 60pt
---

\fancyhead[R]{Josh Pickel \\ STATS419 \\ Instructor: Dr. Shaffer \\ `r format(Sys.time(), '%Y %B %d')`}

```{r, top of the world,echo = FALSE, message= FALSE}
library(devtools)
source_url("https://raw.githubusercontent.com/JoshuaPickel/WSU_STATS419_2021/main/functions/functions-intro/r-intro.R")

```
\par
# HandShake Function

The funtionI created akes in an interger representing the number of people shaking hands, and it determines the number of handshakes that occur. It has an option to plot the results as a network graph, as demonstrated below.


```{r, r-intro handShakefunction, fig.cap = "Network Graph Depicting the Handshakes Between 5 People", message=FALSE}
library(igraph)
handShake(n=10,plotMe = TRUE)

```
With 5 people shaking hands, there is a total of 10 handshakes that can occur. Each handshake is represented in the network graph, showing each person and their respective handshakes. There are arrows on both ends of each line connecting each person as they shook each others hands. 

# Counting Letters From Declaration of Independence Draft and Final Version

The function created for this task takes in the file name of the Declaration of Independence version, and counts the number of letters used in the version being passed. The results are shown below as a table and as a bar chart showing how many times each letter has been used in the file provided to the function.

```{r, letter count for original draft of Declaration of Independence, fig.cap = "Count of Letters Used in original.txt"}
count_letters('original.txt')

```

\newpage
```{r, letter count for final draft of Declaration of Independence,fig.cap = "Count of Letters Used in final.txt"}

count_letters('final.txt')
```

# Matrix Determinant

The function written to compute the determinant of a 3x3 matrix takes in a matrix and checks it is a 3x3. It will then "manually" compute the determinant (by that I mean I did not call on det()). The function is demonstrated below. 

```{r, 3x3 matrix determinant }
myMatrix = matrix(c(6,4,2,1,-2,8,1,5,7),nrow = 4, ncol = 3)
print(myMatrix)
determinant = Determinant(myMatrix)
print(determinant)

```
Using this example matrix "myMatrix", we can see the determinant is -306. 






